Allocation Removal by Partial Evaluation in a Tracing JIT

### 2010-10-14 talk/pepm2011/paper.tex
clarifications

One problem to the successful application of the allocation removal
techniques described in the previous sections is the presence of
frame-introspection features in many dynamic languages. Languages such
as Python and Smalltalk allow the programmer to get access to the
frames object that the interpreter uses to store local variables. This
is a useful feature, as it makes the implementation of a debugger
possible in Python without needing much explicit support from the VM
level. On the other hand, it severely hinders the effectiveness of
allocation removal, because every time an object is stored into a
local variable, it is stored into the frame-object, which makes it
escape.

@@

In this paper, we used an approach based on online partial evaluation to
optimize away allocations in the traces of a tracing JIT.  In this context a
simple approach to partial evaluation gives good results.  This is due
to the fact that the tracing JIT itself is responsible for all control
issues, which are usually the hardest part of partial evaluation: the
tracing JIT selects the parts of the program that are worthwhile to
optimize, and extracts linear paths through them, inlining functions as
necessary.  What is left to optimize is only those linear paths.

>>>>

One problem to the successful application of the allocation removal
techniques described in the previous sections is the presence of
frame-introspection features in many dynamic languages and the fact
that frames are heap allocated. Languages such as Python and Smalltalk
allow the programmer to get access to the frames object that the
interpreter uses to store local variables. This is a useful feature,
as it makes the implementation of a debugger possible in Python
without needing much explicit support from the VM level. On the other
hand, it severely hinders the effectiveness of allocation removal,
because every time an object is stored into a local variable, it is
stored into the frame-object, which makes it escape.

@@

In this paper, we used an approach based on online partial evaluation
to optimize away allocations and type guards in the traces of a
tracing JIT.  In this context a simple approach to partial evaluation
gives good results.  This is due to the fact that the tracing JIT
itself is responsible for all control issues, which are usually the
hardest part of partial evaluation: the tracing JIT selects the parts
of the program that are worthwhile to optimize, and extracts linear
paths through them, inlining functions as necessary.  What is left to
optimize is only those linear paths.
